When, why, and how to avoid, prefer, or tolerate inbreeding

### 2018-03-02 presentation/inbreeding.Rmd
Update slides

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
1. [Wibowo Djatmiko](https://commons.wikimedia.org/wiki/User:Wie146). 2. [George Shuklin](https://commons.wikimedia.org/wiki/User:George_Shuklin). 3. [Natalie Pilakouta](https://www.nataliepilakouta.com/). 4. [Neale Monks](https://en.wikipedia.org/wiki/User:Neale_Monks). 5. [Jane Reid](https://www.abdn.ac.uk/ibes/profiles/jane.reid). 6. [André Karwath](https://commons.wikimedia.org/wiki/User:Aka). 7. [Benjamin Twist](https://commons.wikimedia.org/wiki/User:Benjamint444) 8. [Heinz Seehagel](https://de.wikipedia.org/wiki/Benutzer:HaSee). 
</span></div>

<!---
<hr>
<span style="font-size: 14px;">
1. *Poecilia reticula* image: [Wibowo_Djatmiko](https://commons.wikimedia.org/wiki/User:Wie146) <br>
</span>
--->

@@

@@

## Inbreeding behaviour

- What do we exect to see given inbreeding depression?

- inbreeding avoidance
- inbreeding preference
- inbreeding tolerance

## Inbreeding theory

Describe Parker's model

- Resource and mate limited sexes
- Mate encounter, opportunity to outbreed assumed

## Inbreeding theory

## Inbreeding theory

## Inbreeding theory

## Inbreeding theory

## Inbeeding theory

- Parker's model
- Add a see-saw plot that weighs relatedness versus inbreeding depression

$$\frac{n}{2} (1 + r)(1 - \delta) > \frac{n}{2}$$

Is inclusive fitness from inbreeding (left) greater than inclusive fitness from outbreeding?

## Interpreting Parker's model

Some key takeaways:

- Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression
- Inbreeding depression thresholds differ between females and males causing sexual conflict
- Females should prefer inbreeding below their inbreeding depression threshold (resource limited):
- Males should *tolerate* inbreeding below their threshold, but *never prefer* [1] (sis will be fine)

Puurtinen's paper

## Extensions of Parker's model

>>>>

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
1. [Wibowo Djatmiko](https://commons.wikimedia.org/wiki/User:Wie146). 2. [George Shuklin](https://commons.wikimedia.org/wiki/User:George_Shuklin). 3. [Natalie Pilakouta](https://www.nataliepilakouta.com/). 4. [Neale Monks](https://en.wikipedia.org/wiki/User:Neale_Monks). 5. [Jane Reid](https://www.abdn.ac.uk/ibes/profiles/jane.reid). 6. [André Karwath](https://commons.wikimedia.org/wiki/User:Aka). <br> 7. [Benjamin Twist](https://commons.wikimedia.org/wiki/User:Benjamint444) (benjamint444@gmail.com) 8. [Heinz Seehagel](https://de.wikipedia.org/wiki/Benutzer:HaSee). 9. [Adam Hayward](https://adhayward.wordpress.com/)
</span></div>

## **Avoidance** of breeding with relatives

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Szulkin M, et al. (2013) *Trends Ecol. Evol.* [**28**:205–211](https://doi.org/10.1016/j.tree.2012.10.016).
</span></div>

## **Preference** for breeding with relatives

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Szulkin M, et al. (2013) *Trends Ecol. Evol.* [**28**:205–211](https://doi.org/10.1016/j.tree.2012.10.016).
</span></div>

## No inbreeding avoidance/preference (**Tolerance**)

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Szulkin M, et al. (2013) *Trends Ecol. Evol.* [**28**:205–211](https://doi.org/10.1016/j.tree.2012.10.016).
</span></div>

@@

<center>
![](images/pedigree_9.png)
</center>

<center>
![](images/pedigree_10.png)
</center>

<center>
![](images/pedigree_11.png)
</center>

## Encounter between relatives

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785).
</span></div>

## Encounter between relatives

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785).
</span></div>

## Encounter between relatives

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785).
</span></div>

## Encounter between relatives

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785).
</span></div>

## Encounter between relatives

<center>
![](images/fem_choice_5.png)

</center>

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785).
</span></div>

## Encounter between relatives

<center>
![](images/fem_choice_6.png)

</center>
<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785).
</span></div>

## When to inbreed versus avoid inbreeding?

Is inclusive fitness from avoiding inbreeding (left side) less than inbreeding (right side)?

$$\frac{n}{2} < \frac{n}{2} (1 + r)(1 - \delta)$$

More generally, is inclusive fitness from inbreeding (left) greater than inclusive fitness from outbreeding?

## Key conceptual take-aways

```{r, echo = FALSE, fig.width = 4.5}
ri <- matrix(data=c(1,0.5,0,0.5,1,0,0,0,1),nrow=3,byrow=TRUE);
Inbreed <- function(focal,mate,alt){
  Inb <- ri[focal,focal] + ri[focal,mate];
  Out <- ri[focal,mate]  + ri[focal,alt];
  DEL <- 1 - (Out/Inb);
  return(DEL);
}
rs <- seq(from=0,to=1,by=0.0001);
dl <- rep(x=0,length=length(rs));
for(i in 1:length(rs)){
  ri <- matrix(data=c(1,rs[i],0,rs[i],1,0,0,0,1),nrow=3,byrow=TRUE);
  dl[i] <- Inbreed(focal=1,mate=2,alt=3);
}
par(mar=c(5,5,1,1));
plot(x=rs,y=dl,type="l",lwd=2,
  xlab=expression(paste("Relatedness (",italic(r),")")),
  ylab=expression(paste(italic(delta)," threshold")),
  cex.lab=1.5,cex.axis=1.25,ylim=c(0,1));
Inbreed <- function(focal,mate,alt){
  Inb <- ri[focal,focal] + ri[focal,mate];
  Out <- ri[focal,focal]  + ri[focal,alt];
  DEL <- 1 - (Out/Inb);
  return(DEL);
}
dk <- rep(x=0,length=length(rs));
for(i in 1:length(rs)){
  ri <- matrix(data=c(1,rs[i],0,rs[i],1,0,0,0,1),nrow=3,byrow=TRUE);
  dk[i] <- Inbreed(focal=2,mate=1,alt=3);
}
points(x=rs,y=dk,type="l",lwd=2,
  cex.lab=1.25,cex.axis=1.5,ylim=c(0,1));
polygon(x=c(rs,rev(rs)),y=c(dk,rev(dl)),col="black");
polygon(x=c(rs,rev(rs)),y=c(rep(0,length(dk)),rev(dk)),col="grey40");
points(x=seq(from=0,to=1,by=0.1),y=rep(1,11),type="l");
points(y=seq(from=0,to=1,by=0.1),x=rep(1,11),type="l");
text(x=0.75,y=0.16,labels=expression(paste(italic(Female))),cex=1.75, col = "black");
arrows(x0=0.75,y0=0.2,x1=0.6,y1=0.375, length = 0.15, angle = 30, code = 2, lwd=2);
text(x=0.75,y=0.84,labels=expression(paste(italic(Male))),cex=1.75, col  = "black");
arrows(x0=0.75,y0=0.8,x1=0.6,y1=0.625, length = 0.15, angle = 30, code = 2, lwd=2);
```

- Inbreeding avoidance is predicted given a sufficiently high threshold of inbreeding depression

- Inbreeding depression thresholds differ between females and males causing sexual conflict

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785). <br> Duthie, AB & JM Reid (2015) *PLoS One* [**10**:e0125140](https://bradduthie.github.io/Duthie%26Reid2015.pdf). 
</span></div>

## Key conceptual take-aways

```{r, echo = FALSE, fig.width = 4.5}
ri <- matrix(data=c(1,0.5,0,0.5,1,0,0,0,1),nrow=3,byrow=TRUE);
Inbreed <- function(focal,mate,alt){
  Inb <- ri[focal,focal] + ri[focal,mate];
  Out <- ri[focal,mate]  + ri[focal,alt];
  DEL <- 1 - (Out/Inb);
  return(DEL);
}
rs <- seq(from=0,to=1,by=0.0001);
dl <- rep(x=0,length=length(rs));
for(i in 1:length(rs)){
  ri <- matrix(data=c(1,rs[i],0,rs[i],1,0,0,0,1),nrow=3,byrow=TRUE);
  dl[i] <- Inbreed(focal=1,mate=2,alt=3);
}
par(mar=c(5,5,1,1));
plot(x=rs,y=dl,type="l",lwd=2,
  xlab=expression(paste("Relatedness (",italic(r),")")),
  ylab=expression(paste(italic(delta)," threshold")),
  cex.lab=1.5,cex.axis=1.25,ylim=c(0,1));
Inbreed <- function(focal,mate,alt){
  Inb <- ri[focal,focal] + ri[focal,mate];
  Out <- ri[focal,focal]  + ri[focal,alt];
  DEL <- 1 - (Out/Inb);
  return(DEL);
}
dk <- rep(x=0,length=length(rs));
for(i in 1:length(rs)){
  ri <- matrix(data=c(1,rs[i],0,rs[i],1,0,0,0,1),nrow=3,byrow=TRUE);
  dk[i] <- Inbreed(focal=2,mate=1,alt=3);
}
points(x=rs,y=dk,type="l",lwd=2,
  cex.lab=1.25,cex.axis=1.5,ylim=c(0,1));
polygon(x=c(rs,rev(rs)),y=c(dk,rev(dl)),col="black");
polygon(x=c(rs,rev(rs)),y=c(rep(0,length(dk)),rev(dk)),col="grey40");
points(x=seq(from=0,to=1,by=0.1),y=rep(1,11),type="l");
points(y=seq(from=0,to=1,by=0.1),x=rep(1,11),type="l");
text(x=0.75,y=0.16,labels=expression(paste(italic(Female))),cex=1.75, col = "black");
arrows(x0=0.75,y0=0.2,x1=0.6,y1=0.375, length = 0.15, angle = 30, code = 2, lwd=2);
text(x=0.75,y=0.84,labels=expression(paste(italic(Male))),cex=1.75, col  = "black");
arrows(x0=0.75,y0=0.8,x1=0.6,y1=0.625, length = 0.15, angle = 30, code = 2, lwd=2);
```

- Inbreeding tolerance increases with lower encounter rates (Kokko & Ots 2006)

- Dispersal has effects (Lehmann & Perrin 2003)

- Thresholds change given subsequent mating decisions of relatives (Duthie & Reid 2015)

- Thresholds change given modulation of parental investment (Duthie et al. 2016)

<div align="left" style="line-height: 0.6em;">
<span style="font-size: 14px;">
Parker (1979) *Sexual selection and reproductive competition in insects* (pp. 123–166). Parker (2006) *Philos. Trans. Royal Soc. B.* [**361**:235-259](https://doi.org/10.1098/rstb.2005.1785). <br> Duthie, AB & JM Reid (2015) *PLoS One* [**10**:e0125140](https://bradduthie.github.io/Duthie%26Reid2015.pdf). 
</span></div>

@@
